### 2022-03-15 paper/localpsf.tex
restructuring paper

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space. Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.

We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.

Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$. I.e., 
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x),
\end{equation}
and similar for $v_h$. Replacing $\Aker$ with $\widetilde{\Phi}^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward but tedious algebraic manipulations, yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
\end{equation}
%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following matrix of approximate kernel entries:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \widetilde{\Phi}(p_i, p_j),
\end{equation}
and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
%\begin{equation*}
%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
%\end{equation*}
The finite element matrix representation of our approximation of $\Aop$ is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}

@@

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}). In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.

>>>>

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise. For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$. I.e., 
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}

%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.

%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$. 
Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
\end{equation}
%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the matrix of kernel entries evaluated at pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
\end{equation}
and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
%\begin{equation*}
%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
%\end{equation*}
Our matrix representation of $\Aop$ is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}
Note that $\mathbf{A}$ is not a Galerkin projection of $\Aop$ onto $V_h$ because of interpolation error in the kernel approximation $\Phi^h \approx \Phi$.

%We neglect this interpolation error because it is the same magnitude as the finite element approximation error. 
%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory. Building and storing these matrices is computationally intractible. Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.

@@

The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H \approx \mathbf{\Phi}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}). In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
